\newpage
\phantomsection \label{ack}
\addcontentsline{toc}{chapter}{Acknowledgements}

\chapter*{Acknowledgements}
This dissertation would not have been possible without the support of several wonderful people who helped me throughout my PhD studies. I have to keep this brief, and I hope I have not missed anyone of importance.

Firstly, I want to express my sincere gratitude to my Director of Studies Prof Ruslan Mitkov, from whom I have received invaluable advice and guidance throughout the course of my studies. He is the first person who encouraged me to employ deep learning in my study, and I came a long way following his advice. His insight and knowledge helped to shape the overall direction of the thesis. The continuous opportunities he provided me throughout the course helped me to grow as a researcher. I am very grateful for his continued encouragement and help in spite of his busy schedule.

I want to extend my sincere thanks to my first supervisor Prof Constantin Orăsan for his dedicated guidance, kindness, and support throughout my studies. His vision significantly influenced the direction of the research and the specifics of the technologies used in my experiments. His scientific and technical knowledge immensely benefited my work. Our regular meetings and conversations were always constructive and positive. Above everything, I want to thank him very much for being a friend more than a supervisor. The words can not describe my gratitude to him.

I am particularly grateful to Rocío Caro Quintana, Lucía Bellés Calvera and Encarnación Núñez Ignacio for providing evaluation support for Spanish sentences in translation memory experiments in this thesis. Their support helped me to complete the translation memory section successfully. I would also like to thank Dr María Navas Loro from Universidad Politécnica de Madrid for providing the tool she developed, Añotador, to extract dates from Spanish texts, which helped me to improve the results in translation memory experiments.


The Research Group in Computational Linguistics (RGCL) at the University of Wolverhampton is one of the best places in the world to study computational linguistics. I had the pleasure of meeting and forming friendships with many talented people from RGCL during this time. The collaborations I had with them helped to improve my technical skills. I am particularly grateful to all of my co-authors, Dr Marcos Zampieri, Alistair Plum, Dr Pablo Calleja, Dr Emad Mohamed, Zeses Pitenis, Hadeel Saadany and Salim Al Mandhari. I would also like to thank Dr Richard Evans, Dr Le An Ha, Dr Frédéric Blain, Dr Michael Oakes, Dr Burcu Can, Dr Omid Rohanian, Dr Maria Kunilovskaya, Dr Najah Albaqawi, Dr Sara Može and Dr Shiva Taslimipoor for having relaxing and informative discussions that helped me to improve my research at various points in my PhD. My time in England is my first time away from home. I want to thank Alistair Plum, Antonio Pascucci, Dr Marcos Zampieri and Dr Pablo Calleja for their friendship and support, which helped me settle down quickly in England.  

A special ‘thank you’ to Kate Wilson for proofreading the thesis and various research papers and providing English support which immensely helped me in writing my PhD. I would also like to thank Dr Amanda Bloore, Dr Emma Franklin, April Harper and Suman Hira for administrative support and proofreading papers at various points during my PhD. Their support made it possible for me to attend various conferences and complete my PhD smoothly.

It has been an honour to have Prof Lynne Bowker and Prof Michael Thelwall as my examiners. I appreciate their valuable comments and the time they spent to read and evaluate the thesis. They were very fast with the reviewing process. I would also like to thank the chair of my viva, Dr Benjamin Halligan for smoothly handling the whole process.

None of these would be possible without emotional support. I want to thank my wife, Hansi Hettiarachchi, for being with me through thick and thin and for the various support she gave me throughout the course, including reading this thesis multiple times. Finally, I would like to thank my mother, father and sister for their unconditional love, blessings and support during the whole PhD.



